A Mechanical Formalization of Higher-Ranked Polymorphic Type Inference

### 2019-01-24 sections/Overview.tex
Declarative typing rules

\begin{figure}[t]
\[
\begin{array}{l@{\qquad}lcl}
\text{Type variables}\qquad&a, b\\[3mm]
\text{Types}\qquad&A, B, C &::=&\quad 1 \mid a \mid \forall a. A \mid A\to B\\
\text{Monotypes}\qquad&\tau &::=&\quad 1 \mid a \mid \tau_1\to \tau_2\\
\text{Contexts}\qquad&\Psi &::=&\quad \cdot \mid \Psi, a\\
%\text{Judgments}\qquad&\exps &::=&\quad \cdot \mid A \le B : \exps
\end{array}
\]
\caption{Syntax of Declarative System}\label{fig:decl:syntax}
\end{figure}

@@

\begin{figure}[t]
\centering \framebox{$\Om \vdash A$}
\begin{gather*}
\inferrule*[right=$\mathtt{wf_d unit}$]
    {~}{\Om\vdash 1}
\qquad
\inferrule*[right=$\mathtt{wf_d var}$]
    {a\in\Om}{\Om\vdash a}
\qquad
\inferrule*[right=$\mathtt{wf_d{\to}}$]
    {\Om\vdash A\quad \Om\vdash B}
    {\Om\vdash A\to B}
\qquad
\inferrule*[right=$\mathtt{wf_d\forall}$]
    {\Om, a\vdash A}
    {\Om\vdash \forall a. A}
\end{gather*}

\centering \framebox{$\Om \vdash A \le B$}
\begin{gather*}
\inferrule*[right=$\mathtt{{\le}Var}$]
    {a\in\Om}{\Om\vdash a\le a}
\qquad
\inferrule*[right=$\mathtt{{\le}Unit}$]
    {~}{\Om \vdash 1 \le 1}
\qquad
\inferrule*[right=$\mathtt{{\le}{\to}}$]
    {\Om \vdash B_1 \le A_1 \quad \Om \vdash A_2 \le B_2}
    {\Om\vdash A_1\to A_2 \le B_1\to B_2}
\\
\inferrule*[right=$\mathtt{{\le}\forall L}$]
    {\Om\vdash \tau \quad \Om\vdash [\tau/a] A \le B}
    {\Om\vdash \forall a. A \le B}
\qquad
\inferrule*[right=$\mathtt{{\le}\forall R}$]
    {\Om, a\vdash A\le B}
    {\Om\vdash A \le \forall a. B}
\end{gather*}
\caption{Well-formedness of Declarative Types and Declarative Subtyping}\label{fig:decl:wf_sub}
\end{figure}

In Figure~\ref{fig:decl:wf_sub},
we give the well-formedness and subtyping relation for the declarative system.

>>>>

\begin{figure}[t]
\[
\begin{array}{l@{\qquad}lcl}
\text{Type variables}\qquad&a, b\\[3mm]
\text{Types}\qquad&A, B, C &::=&\quad 1 \mid a \mid \forall a. A \mid A\to B\\
\text{Monotypes}\qquad&\tau,\sigma &::=&\quad 1 \mid a \mid \tau\to \sigma\\
\text{Contexts}\qquad&\Psi &::=&\quad \cdot \mid \Psi, a\\
%\text{Judgments}\qquad&\exps &::=&\quad \cdot \mid A \le B : \exps
\end{array}
\]
\caption{Syntax of Declarative System}\label{fig:decl:syntax}
\end{figure}

@@

\begin{figure}[t]
%\centering \framebox{$\Psi \vdash A$}
%\begin{gather*}
%\inferrule*[right=$\mathtt{wf_d unit}$]
%    {~}{\Psi\vdash 1}
%\qquad
%\inferrule*[right=$\mathtt{wf_d var}$]
%    {a\in\Psi}{\Psi\vdash a}
%\qquad
%\inferrule*[right=$\mathtt{wf_d{\to}}$]
%    {\Psi\vdash A\quad \Psi\vdash B}
%    {\Psi\vdash A\to B}
%\qquad
%\inferrule*[right=$\mathtt{wf_d\forall}$]
%    {\Psi, a\vdash A}
%    {\Psi\vdash \forall a. A}
%\end{gather*}

\centering \framebox{$\Psi \vdash A \le B$}
\begin{gather*}
\inferrule*[right=$\mathtt{{\le}Var}$]
    {a\in\Psi}{\Psi\vdash a\le a}
\qquad
\inferrule*[right=$\mathtt{{\le}Unit}$]
    {~}{\Psi \vdash 1 \le 1}
\qquad
\inferrule*[right=$\mathtt{{\le}{\to}}$]
    {\Psi \vdash B_1 \le A_1 \quad \Psi \vdash A_2 \le B_2}
    {\Psi\vdash A_1\to A_2 \le B_1\to B_2}
\\
\inferrule*[right=$\mathtt{{\le}\forall L}$]
    {\Psi\vdash \tau \quad \Psi\vdash [\tau/a] A \le B}
    {\Psi\vdash \forall a. A \le B}
\qquad
\inferrule*[right=$\mathtt{{\le}\forall R}$]
    {\Psi, a\vdash A\le B}
    {\Psi\vdash A \le \forall a. B}
\end{gather*}
\caption{%Well-formedness of Declarative Types and 
Declarative Subtyping}\label{fig:decl:sub}
\end{figure}

\begin{figure}[t]
\centering \framebox{$\Psi \vdash A \Lto B$} $e$ checks against input type $A$.\\
\centering \framebox{$\Psi \vdash A \To B$} $e$ synthesizes output type $A$.\\
\centering \framebox{$\Psi \vdash \appInf{A}{e}{C}$} Applying a function of type $A$ to $e$ synthesizes type $C$.
\begin{gather*}
\inferrule*[right=$\mathtt{DeclVar}$]
    {(x:A)\in\Psi}{\Psi\vdash x\To A}
\qquad
\inferrule*[right=$\mathtt{DeclSub}$]
    {\Psi\vdash e\To A \quad \Psi\vdash A\le B}{\Psi \vdash e\Lto B}
\qquad
\inferrule*[right=$\mathtt{DeclAnno}$]
    {\Psi \vdash A \quad \Psi\vdash e\Lto A}
    {\Psi\vdash (e:A)\To A}
\\
\inferrule*[right=$\mathtt{Decl1I}$]
    {~}{\Psi\vdash () \Lto 1}
\qquad
\inferrule*[right=$\mathtt{Decl1I\To}$]
    {~}{\Psi\vdash () \To 1}
\qquad
\inferrule*[right=$\mathtt{Decl\forall I}$]
    {\Psi,a \vdash e \Lto A}
    {\Psi\vdash e\Lto \all A}
\\
\inferrule*[right=$\mathtt{Decl\forall App}$]
    {\Psi \vdash \tau \quad \Psi\vdash \appInf{[\tau/a]A}{e}{C} }
    {\Psi\vdash \appInf{\all A}{e}{C}}
\qquad
\inferrule*[right=$\mathtt{Decl\to I}$]
    {\Psi,x:A \vdash e\Lto B}
    {\Psi\vdash \lam e \Lto A \to B}
\\
\inferrule*[right=$\mathtt{Decl\to I\To}$]
    {\Psi\vdash \sigma\to\tau \quad \Psi,x:\sigma \vdash e\Lto \tau}
    {\Psi\vdash \lam e \To \sigma\to\tau}
\qquad
\inferrule*[right=$\mathtt{Decl\to E}$]
    {\Psi\vdash e_1\To A \quad \Psi\vdash \appInf{A}{e_2}{C}}
    {\Psi\vdash e_1~e_2 \To C}
\\
\inferrule*[right=$\mathtt{Decl\to App}$]
    {\Psi\vdash e \Lto A}
    {\Psi\vdash \appInf{A \to C}{e}{C}}
\end{gather*}
\caption{Declarative Typing}\label{fig:decl:typing}
\end{figure}

In Figures~\ref{fig:decl:sub} and \ref{fig:decl:typing},
we give the subtyping and typing relations for the declarative system.
